### 2019-11-01 defesa/tex/chapter/e3p.tex
wip

According to \citeonline[p.~150]{powell}, any function of the form $\{\cos^j{x}\sin^k{x} : j, k \in \mathbb{N}\}$ can be written as a real trigonometric polynomial of degree $j+k$ which can have up to $2(j+k)$ different roots in the interval $[0, 2\pi)$. The reason to bring up this fact is that $\xi(\theta)$ can be written as $\sum_i^M c_i \cos^{j_i}(\theta)\sin^{k_i}(\theta)$, for some $M \in \mathbb{N}$ and $c_1, \dots, c_m \in \R$, which implies that it can have up to $2$ times its degree.

 To show that, just note that it is possible to write $\norm{\varphi(v, \theta)}^2$ and $A(\theta)^2$ in that form, as it can be seen on \autoref{eq:dd} and \autoref{eq:dd2}, combine the parts as $\xi(\theta)$ is just addition and multiplication of terms in that form. It is also possible to see that the term of higher the degree of $\xi(\theta)$ is the multiplication of the three squared distances, as $\norm{\varphi(v, \theta)}^2$ has degree $2$ the degree of $\xi(\theta)$ is $6$.

\begin{align}\label{eq:dd}
	\norm{\varphi(v, \theta)}^2 = (v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta)^2 + (v_y\cos\theta - v_x\sin\theta)^2\\
	\label{eq:dd2} A(\theta)^2=\dfrac{1}{4}\det\left(
	\begin{array}{cc}
		v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta&v_y\cos\theta - v_x\sin\theta\\
		w_x\frac{b}{a}\cos\theta + w_y\frac{b}{a}\sin\theta&w_y\cos\theta - w_x\sin\theta
	\end{array}\right)^2
\end{align}

@@

\subsection{Converting $\xi(\theta)$ into a polynomial}

Using the identity $x = \tan{\frac{\theta}{2}}$, it is possible to convert $\xi(\theta)$ on \autoref{eq:circumscribed_circle_b} into a univariate polynomial of degree $12$. The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. Therefore, pursuing a numerical method would be a good choice for this problem.

@@

This method seemed to work fine, but for some instances it did not find roots which were priorly known. This happened probably because of the transformation that was used to convert $\xi$ into a polynomial. It can be very numerically unstable as the tangent function grows to fast for angles that are greater than $\frac{\pi}{4}$. Depending on the instance, there could be roots very close to $0$ and others that are too far from it.

@@

\section{A method for E3PNT}

@@

\subsection{Chebyshev interpolation}

@@

\begin{equation}\label{eq:lagrange}
f_n(x) = \sum_{j=0}^{n} f(x_j)\dfrac{\prod_{k \neq j}^{n+1} (x-x_k)}{\prod_{k \neq j}^{n+1} (x_j-x_k)},
\end{equation} 
with $f$ being the function to be approximated and $f_n$ the unique $n$ degree polynomial that passes through $\{(x_j, f(x_j)): j=0, 1, \dots n\}$. Because of the uniqueness of interpolant polynomials, there is a direct link between the quality of an approximation and the points chosen to interpolate. As a matter of fact, depending on the points one chooses, even increasing the degree of the interpolation makes the approximation worsen. This is known as Runge's phenomenon and an example can be seen in \citeonline[p.~37]{powell} where uniformly spaced points are chosen to interpolate the function $f(x) = (1+x^2)^{-1}$ on the interval $[-5, 5]$. That is where Chebyshev interpolation comes in, instead of $n+1$ arbitrary points, the interpolation is done on the $n+1$ roots of $T_n$, which are also known as Chebyshev Nodes:

\begin{equation}
x_j = \cos{\left(\pi\dfrac{2n-2j+1}{2n}\right)},
\end{equation}
for $j=1, \dots, n$. If the domain of the function to be interpolated is in another range other than $[-1, 1]$, let us say $[a, b]$, then a transformation can be done to map it to the Chebyshev Nodes' domain:

\begin{equation}
\hat{x_j} = \frac{a+b}{2} + \frac{b-a}{2}x_j.
\end{equation}

Then, the Chebyshev interpolation of a function $f: [a, b] \mapsto \R$ can be determined using Lagrange's formula and the points $\hat{x}_1, \dots, \hat{x}_n$. 
As it was mentioned, finding the roots of a polynomial written in the monomial form can be done by determining the eigenvalues of a so-called Frobenius companion matrix. For small $n$ this works fine, however, converting the polynomial obtained by \autoref{eq:lagrange} to the power form, as $n$ grows, becomes a very ill-conditioned problem.

An alternative method can be found in \citeonline{boyd:2013} where the Chebyshev interpolation is calculated directly as a truncated Chebyshev series, as in \autoref{eq:chebseries}, in $\bigO(n^2)$. Also, given a polynomial written in the Chebyshev basis, a $n\times n$ matrix can be constructed, such that its eigenvalues are the roots of that polynomial. \citeonline{boyd:2013} refers to this matrix as the Chebyshev-Frobenius companion matrix.

@@

Even though $\xi(\theta)$ is complicated enough, in a sense that finding its roots directly is no trivial task, it is a very well-behaved function: it is analytic and  has infinitely many continuous and integrable derivatives. This satisfy all the requirements of the result in \citeonline[p.~28]{gottlieb} which says that if a function has $m$ continuous and integrable derivatives on a closed interval, then its absolute difference between the Chebyshev truncate series is $\bigO(n^{-m})$. Also, in \citeonline{battles:2004} a theorem is presented stating that if a function is analytic on a neighborhood of $[-1, 1]$, then the convergence is $\bigO(C^n)$, for some $C<1$.

@@

\begin{equation}
r_n = \max_{0 \le \theta \le \pi} |f_n(\theta) - \xi(\theta)|.
\end{equation}

>>>>

According to \citeonline[p.~150]{powell}, any function of the form $\{\cos^j{x}\sin^k{x} : j, k \in \mathbb{N}\}$ can be written as a real trigonometric polynomial of degree $j+k$ which can have up to $2(j+k)$ different roots in the interval $[0, 2\pi)$. The reason to bring up this fact is that $\xi$ can be written as $$\sum_{0 \le j+k \le 6} c_{j,k} \cos^{j}(\theta)\sin^{k}(\theta),$$ for some $\{c_{j,k} \in \R: 0 \le j+k \le 6\}$, which implies that it can be converted to a real trigonometric polynomial of degree $6$ which can have up to $12$ roots.

 To show that, just note that it is possible to write $\norm{\varphi(v, \theta)}^2$ and $A(\theta)^2$ in that form, as it can be seen on \autoref{eq:dd} and \autoref{eq:dd2}. It is also possible to see that the term of higher the degree of $\xi$ is the multiplication of the three squared distances, as $\norm{\varphi(v, \theta)}^2$ has degree $2$, the degree of $\xi$ is $6$.
\begin{align}\label{eq:dd}
	\norm{\varphi(v, \theta)}^2 = (v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta)^2 + (v_y\cos\theta - v_x\sin\theta)^2\\
	\label{eq:dd2} A(\theta)^2=\dfrac{1}{4}\det\left(
	\begin{array}{cc}
		v_x\frac{b}{a}\cos\theta + v_y\frac{b}{a}\sin\theta&v_y\cos\theta - v_x\sin\theta\\
		w_x\frac{b}{a}\cos\theta + w_y\frac{b}{a}\sin\theta&w_y\cos\theta - w_x\sin\theta
	\end{array}\right)^2
\end{align}

@@

\subsection{Converting $\xi$ into a polynomial}

Using the identity $x = \tan{\frac{\theta}{2}}$, it is possible to convert $\xi(\theta)$ on \autoref{eq:circumscribed_circle_b} into a univariate real polynomial of degree $12$. The famous Abel-Ruffini Theorem (a proof can be seen in \citeonline{skopenkov2015}) states that for polynomials of degree higher than four, there is no closed formula to determine their roots. Therefore, pursuing a numerical method would be a good choice for this problem.

@@

This method seemed to work fine, but for some instances it did not find roots which were priorly known. This happened probably because of the transformation that was used to convert $\xi$ into a polynomial--later we introduce a different transformation which is more convenient. It can be very numerically unstable as the tangent function grows to fast for angles that are greater than $\frac{\pi}{4}$. Depending on the instance, there could be roots very close to $0$ and others that are too far from it.

@@

\section{Two methods for E3PNT}

In this section we introduce two methods that successfully find every solution of E3P on the instances we use for experiments. 
The first one uses an approximation of $\xi$, it is slightly more efficient in respect to running time than the second method, however, it the approximation has only asymptotic guarantees and we can assure that it will always work. The second method converts $\xi$ into a complex polynomial and finds its roots by determining the eigenvalues of its companion matrix.

\subsection{An approximation method}

@@

\subsubsection{Chebyshev interpolation}

@@

\begin{equation}\label{eq:lagrange}
f_n(x) = \sum_{j=0}^{n} f(x_j)\dfrac{\prod_{k \neq j}^{n+1} (x-x_k)}{\prod_{k \neq j}^{n+1} (x_j-x_k)},
\end{equation} 
with $f$ being the function to be approximated, and $f_n$ the unique $n$-degree polynomial that passes through $\{(x_j, f(x_j)): j=0, 1, \dots n\}$. Because of the uniqueness of interpolant polynomials, there is a direct link between the quality of an approximation and the points chosen to interpolate. As a matter of fact, depending on the points one chooses, even increasing the degree of the interpolation makes the approximation worsen. This is known as Runge's phenomenon and an example can be seen in \citeonline[p.~37]{powell} where uniformly spaced points are chosen to interpolate the function $f(x) = (1+x^2)^{-1}$ on the interval $[-5, 5]$.

That is where Chebyshev interpolation comes in. Instead of choosing $n+1$ arbitrary points, the $n+1$ roots of $T_{n+1}$, which are also known as Chebyshev Nodes, are chosen as the interpolation points:
\begin{equation}
x_j = \cos{\left(\dfrac{\pi(k-\frac{1}{2})}{n+1}\right)},
\end{equation}
for $j=1, \dots, n+1$. This particular choice defeats Runge's phenomenon and provides a convergent approximation.

Note that, if the domain of the function to be interpolated is defined on a range other than $[-1, 1]$, let us say $[a, b]$, then a transformation can be done to map it to the Chebyshev Nodes' domain:
\begin{equation}
\hat{x_j} = \frac{a+b}{2} + \frac{b-a}{2}x_j.
\end{equation}

Then, the Chebyshev interpolation of a function $f: [a, b] \mapsto \R$ can be determined using Lagrange's formula and the points $\hat{x}_1, \dots, \hat{x}_n$. 
As it was mentioned, finding the roots of a polynomial written in the monomial form can be done by determining the eigenvalues of a so-called Frobenius companion matrix. For small $n$ this works fine, however, converting the polynomial obtained by \autoref{eq:lagrange} to the power form, as $n$ grows, becomes a very ill-conditioned problem. 
An alternative method can be found in \citeonline{boyd:2013} where the Chebyshev interpolation is calculated directly as a truncated Chebyshev series, as in \autoref{eq:chebseries}, in $\bigO(n^2)$. Also, given a polynomial written in the Chebyshev basis, a $n\times n$ matrix can be constructed, such that its eigenvalues are the roots of that polynomial. \citeonline{boyd:2013} refers to this matrix as the Chebyshev-Frobenius companion matrix.

@@

Even though $\xi$ is complicated enough, in a sense that finding its roots directly is no trivial task, it is a very well-behaved function: it is analytic and  has infinitely many continuous and integrable derivatives. This satisfy all the requirements of the result in \citeonline[p.~28]{gottlieb} which says that if a function has $m$ continuous and integrable derivatives on a closed interval, then its absolute difference between the Chebyshev truncate series is $\bigO(n^{-m})$. Also, in \citeonline{battles:2004} a theorem is presented stating that if a function is analytic on a neighborhood of $[-1, 1]$, then the convergence is $\bigO(C^n)$, for some $C<1$.

To choose the degree of the interpolation we use the last coefficient rule-of-thumb introduced by \citeonline[p.~50]{boyd:2001}. There is no guarantee that this method will choose $n$, such that $f_n$ is close enough to $\xi$ everywhere on $[0, \pi)$, nonetheless, in practice, it is taken to be a good estimate for the error $r_n$:
\begin{equation}
r_n = \max_{0 \le \theta \le \pi} |f_n(\theta) - \xi(\theta)|.
\end{equation}

\subsection{Transforming $\xi$ into a complex polynomial}

This approach is based on a idea published on \citeonline{boyd:2006} which converts a real trigonometric polynomial into a polynomial with a variable $z = e^{i\theta}$. In our case, we replace $\cos{\theta}$ and $\sin{\theta}$ with the following identities:
\begin{align}
\cos{\theta} = \dfrac{e^{i\theta} + e^{-i\theta}}{2}\\
\sin{\theta} = \dfrac{e^{i\theta} - e^{-i\theta}}{2i}.
\end{align}

It is possible to show that with that substitution and changing the variable to $z=e^{it}$, we obtain a function on the form:
\begin{equation}
g(z)=\sum_{k=-6}^6 c_k z^k.
\end{equation}
Then, multiplying it by

@@
